Applications of limited information strategies in Menger's game

### 2015-08-26 articles/applications-of-limited-information-strategies-in-mengers-game/article.tex
finish draft sans abstract and introduction

\begin{definition}
  For two functions \(f,g\) we say \(f\) is \textbf{\(\mu\)-almost compatible} with
  \(g\) (\(f\alcomp_\mu g\)) if \(|\{x\in\dom(f)\cap\dom(g):f(x)\not=g(x)\}|<\mu\).
  If \(\mu=\omega\) then we say \(f,g\) are \textbf{almost compatible}
  (\(f\alcomp g\)).
\end{definition}

@@

\begin{theorem}
  \(\alcompS{\omega_1}\) and \(\kappa>2^\omega\Rightarrow\neg\alcompS\kappa\)
  are theorems of \(ZFC\).
  \(\alcompS{2^\omega}\) is a theorem of \(ZFC+CH\) and consistent with
  \(ZFC+\neg CH\).
\end{theorem}

\begin{proof}
  For \(\alcompS{\omega_1}\), look at pg. 70 of \cite{MR597342}; this of course
  implies \(\alcompS{2^\omega}\) under \(CH\).
  \(\neg\alcompS{(2^\omega)^+}\) is shown by a cardinality argument
  in \cite{MR1129143}.
  The consistency result under \(ZFC+\neg CH\)
  is a lemma for the main theorem in \cite{MR1129143}.
\end{proof}

@@

Note that while it's consistent that
\(\pl F\kmarkwin2\menGame{\oneptlind{(2^\omega)}}\), \(\oneptlind{\kappa}\)
for \(\kappa>2^\omega\) is a candidate to answer the above question.

We are also interested in non-game-theoretic characterizations of these
covering properties. It has been known for some time that for metrizable spaces,
winning Menger spaces are exactly the \(\sigma\)-compact spaces, shown first
by Telgarksy in \cite{MR753073} and later directly by Scheepers in
\cite{MR1273523}.

@@

For example, any bounded subset of Euclidean space is relatively compact
whether it is closed or not. Actually, relative compactness can be thought
of as an analogue of boundedness for regular spaces.

\begin{proposition}
  For regular spaces, \(Y\) is relatively compact to \(X\) if and only if
  \(\cl Y\) is compact in \(X\).
  \footnote{
    It should be noted that some authors define relative compactness in
    this way, but such a definition creates pathological implications for
    non-regular spaces. For example, the singleton containing
    the particular point of an infinite space with the particular point topology
    would not be relatively compact since its closure is not compact, even
    though it is finite.
  }
\end{proposition}

\begin{proof}
  For any space, any subset of a compact set is relatively compact.

  Assume \(Y\) is relatively compact, let \(\mc U\) be an open cover of \(\cl Y\),
  and define \(x\in V_x\subseteq \cl{V_x}\subseteq U_x \in \mc U\) for each
  \(x\in X\). Then if we take a subcollection \(\mc F = \{V_{x_i} : i<n\}\) covering
  \(Y\) by relative compactness, then \(\{U_{x_i}:i<n\}\) is a finite subcollection
  of \(\mc U\) covering \(\cl Y\), showing compactness.
\end{proof}

We now begin the process of factoring out Scheeper's proof to reveal the
limited information implications at work.

@@

\begin{corollary}
  The following are equivalent:
  \begin{itemize}
    \item \(X\) is \(\sigma\)-relatively-compact
    \item \(\pl F \prewin \menGame X\)
    \item \(\pl F \markwin \menGame X\)
  \end{itemize}
\end{corollary}

\begin{proof}
  If \(X=\bigcup_{n<\omega} R_n\) for \(R_n\) relatively compact, then
  \(\sigma(n)=R_n\) is a winning predetermined strategy, which yields a
  winning Markov strategy. The previous lemma finishes the proof.
\end{proof}

\begin{corollary}
  Let \(X\) be a regular space. The following are equivalent:
  \begin{itemize}
    \item \(X\) is \(\sigma\)-compact
    \item \(X\) is \(\sigma\)-relatively-compact
    \item \(\pl F \prewin \menGame X\)
    \item \(\pl F \markwin \menGame X\)
  \end{itemize}
\end{corollary}

@@

\begin{corollary}
  Let \(X\) be a second-countable space. The following are equivalent:
  \begin{itemize}
    \item \(X\) is \(\sigma\)-relatively-compact
    \item \(F \prewin \menGame{X}\)
    \item \(F \markwin \menGame{X}\)
    \item \(F \win \menGame{X}\)
  \end{itemize}
\end{corollary}

\begin{corollary}
  Let \(X\) be a metrizable space. The following are equivalent:
  \begin{itemize}
    \item \(X\) is \(\sigma\)-compact
    \item \(X\) is \(\sigma\)-relatively-compact
    \item \(F \prewin \menGame{X}\)
    \item \(F \markwin \menGame{X}\)
    \item \(F \win \menGame{X}\)
  \end{itemize}
\end{corollary}

\begin{proof}
  Each property implies Lindel\"of, so \(X\) may be assumed to be
  regular and second-countable.
\end{proof}

@@

To help describe \(\pl F\kmarkwin2\menGame{X}\) topologically, we introduce a
subset variant of the Menger game and a related covering property.

@@

We may also consider common (non-regular) counterexamples which are finer
than the usual Euclidean line.

@@

\begin{theorem}
  \(R_{\mathbb Q}\) is non-regular and non-\(\sigma\)-compact, but is
  second-countable and \(\sigma\)-relatively-compact.
\end{theorem}

\begin{proof}
  Compact sets in \(R_{\mathbb Q}\) can be shown to not contain open intervals,
  and thus are nowhere dense in nonmeager \(\mathbb R\),
  so \(R_{\mathbb Q}\) is not \(\sigma\)-compact. The usual base of
  intervals with rational endpoints (with or without rationals removed)
  witnesses second-countability.

  To see that \(R_{\mathbb Q}\) is \(\sigma\)-relatively compact, consider
  \([a,b]\setminus\mathbb{Q}\). Let \(\mc U\) be a cover of \(R_{\mathbb Q}\), and
  let \(\mc U'\) fill in the missing rationals for any open set in \(\mc U\).
  There is a finite subcover \(\mc V'\subseteq \mc U'\) for \([a,b]\) since
  \(\mc U'\) contains open sets from the Euclidean topology. Let
  \(\mc V= \{V\setminus\mathbb{Q}:V\in V'\}\): this is a finite refinment of
  \(\mc U\) covering \([a,b]\setminus\mathbb{Q}\), so \([a,b]\setminus\mathbb{Q}\)
  is relatively compact. It follows then that \(R_{\mathbb Q}\setminus\mathbb Q\)
  is \(\sigma\)-relatively-compact, and since \(\mathbb Q\) is countable,
  \(R_{\mathbb Q}\) is \(\sigma\)-relatively-compact. Non-regularity follows since
  regular and \(\sigma\)-relatively-compact implies \(\sigma\)-compact.
\end{proof}

@@

\begin{theorem}
  \(R_\omega\) is non-regular, non-second-countable, and
  non-\(\sigma\)-relatively-compact, but \(\pl F\win\menGame{R_\omega}\).
\end{theorem}

\begin{proof}
  The closure of any open set is its closure in the usual Euclidean topology,
  so \(R_\omega\) is not regular. If \(S\supseteq\{s_n:n<\omega\}\) for \(s_n\)
  discrete, then \(U_m=R_\omega\setminus\{s_n:m<n<\omega\}\) yields an
  infinite cover \(\{U_m:m<\omega\}\) with no finite subcollection covering \(S\),
  showing that all relatively compact sets are finite, and \(R_\omega\) is not
  \(\sigma\)-relatively-compact.

  Define the winning strategy \(\sigma\) for \(\pl F\) in \(\menGame{R_\omega}\) as
  follows: let \(\sigma(\mc{U}_0,\dots,\mc{U}_{2n})=[-n,n]\setminus C_n\)
  for some countable \(C_n=\{c_{n,m}: m<\omega\}\),
  and let \(\sigma(\mc{U}_0,\dots,\mc{U}_{2n+1})=\{c_{i,j} : i,j < n\}\).
  Non-second-countable follows since
  second-countable and \(\pl F\win\menGame{X}\) implies
  \(\sigma\)-relatively-compact.
\end{proof}

We will soon see that, assuming \(\alcompS{2^\omega}\), \(\pl F\) has a winning
\(2\)-Mark\"ov strategy for \(\menGame{R_\omega}\) as well.

\begin{proposition}
  Let \(\limitwin\) be either \(\kmarkwin{k}\) or \(\win\).
  If \(X=\bigcup_{i<\omega} X_i\) and \(\pl F \limitwin \menGame{X,X_i}\)
  for \(i<\omega\), then \(\pl F \limitwin \menGame{X}\)
\end{proposition}

\begin{proof}
  Let \(L\) be the \(k\)-Markov fog-of-war \(\mu_k\) (resp. the identity),
  and let \(\sigma_i\) be a \(k\)-Markov strategy (resp. perfect information
  strategy) for \(\pl F\) in \(\menGame{X,X_i}\).

  We define the \(k\)-Markov strategy (resp. perfect information strategy)
  \(\sigma\) for \(\menGame{X}\) as follows:
    \[
      \sigma\circ L(\<\mc U_0,\dots,\mc U_n\>)
        =
      \bigcup_{i\leq n}
      \sigma_i\circ L(\<\mc U_i,\dots,\mc U_n\>)
    \]

  Let \(\<\mc U_0,\mc U_1,\dots\>\) be a successful counter-attack by \(\pl C\)
  against \(\sigma\). Then there exists \(x\in X_i\) for
  some \(i<\omega\) such that \(x\) is not covered by
  \(\bigcup_{n<\omega}\sigma\circ L(\<\mc U_0,\dots,\mc U_n\>)\).
  It follows that \(x\) is not covered by
  \(\bigcup_{n<\omega}\sigma_i\circ L(\<\mc U_i,\dots,\mc U_{i+n}\>)\),
  and \(\<\mc U_i,\mc U_{i+1},\dots\>\) is a successful counter-attack
  by \(\pl C\) against \(\sigma_i\).
\end{proof}

\begin{theorem}
  If \(\alcompS{2^\omega}\), then \(\pl F\kmarkwin2\menGame{R_\omega}\).
\end{theorem}

\begin{proof}
  It's sufficient to show that \([0,1]\subseteq R_\omega\) is relatively
  robustly Menger. Let \(f_A\) witness \(\alcompS{2^\omega}\)
  for \(A\in[[a,b]]^{\leq\omega}\).
  For each open cover \(\mc U\), let \(A_{\mc U}\) be
  such that \([0,1]\setminus A_{\mc U}\) is \(\mc U\)-finite.
  Let \(r_{\mc U}(x)=0\) if \(x\in[0,1]\setminus A_{\mc U}\) and
  \(r_{\mc U}(x)=f_{\mc A_{\mc U}}(x)\) otherwise.

>>>>

\begin{definition}
  For two functions \(f,g\) we say \(f\) is \textbf{almost compatible} with
  \(g\) (\(f\alcomp g\)) if \(|\{x\in\dom(f)\cap\dom(g):f(x)\not=g(x)\}|<\omega\).
\end{definition}

@@

\begin{theorem}
  \(\alcompS{\omega_1}\) and \(\neg\alcompS{\mf c^+}\)
  are theorems of \(ZFC\).
  \(\alcompS{\mf c}\) is a theorem of \(ZFC+CH\) and consistent with
  \(ZFC+\neg CH\).
\end{theorem}

\begin{proof}
  For \(\alcompS{\omega_1}\), consider the construction of an Aronzajn tree
  as in \cite{MR597342}; this of course
  implies \(\alcompS{\mf c}\) under \(CH\).
  \(\neg\alcompS{\mf c^+}\) is shown by a cardinality argument
  in \cite{MR1129143}.
  The consistency result under \(ZFC+\neg CH\)
  is a lemma for the main theorem in \cite{MR1129143}.
\end{proof}

@@

Note that while it's consistent that
\(\pl F\kmarkwin2\menGame{\oneptlind{\mf c}}\), \(\oneptlind{\kappa}\)
for \(\kappa>\mf c\) is a candidate to answer the above question.

We are also interested in non-game-theoretic characterizations of these
covering properties. It has been known for some time that metrizable perfect
Menger spaces are exactly the metrizable \(\sigma\)-compact spaces, shown first
by Telgarksy in \cite{MR753073} and later directly by Scheepers in
\cite{MR1273523}.

@@

\begin{proposition}
  For regular spaces, \(Y\) is relatively compact to \(X\) if and only if
  \(\cl Y\) is compact in \(X\).
\end{proposition}

@@

\begin{corollary}
  The following are equivalent:
  \begin{itemize}
    \item \(X\) is \(\sigma\)-relatively-compact
    \item \(\pl F \kmarkwin0 \menGame X\)
    \item \(\pl F \markwin \menGame X\)
  \end{itemize}
\end{corollary}

\begin{proof}
  If \(X=\bigcup_{n<\omega} R_n\) for \(R_n\) relatively compact, then
  \(\sigma(n)=R_n\) is a winning \(0\)-mark, which of course gives a
  winning \(1\)-mark. The previous lemma finishes the proof.
\end{proof}

\begin{corollary}
  Let \(X\) be a regular space. The following are equivalent:
  \begin{itemize}
    \item \(X\) is \(\sigma\)-compact
    \item \(X\) is \(\sigma\)-relatively-compact
    \item \(\pl F \kmarkwin0 \menGame X\)
    \item \(\pl F \markwin \menGame X\)
  \end{itemize}
\end{corollary}

@@

\begin{corollary}
  Let \(X\) be a second-countable space. The following are equivalent:
  \begin{itemize}
    \item \(X\) is \(\sigma\)-relatively-compact
    \item \(F \kmarkwin0 \menGame{X}\)
    \item \(F \markwin \menGame{X}\)
    \item \(F \win \menGame{X}\)
  \end{itemize}
\end{corollary}

\begin{corollary}
  Let \(X\) be a metrizable space. The following are equivalent:
  \begin{itemize}
    \item \(X\) is \(\sigma\)-compact
    \item \(X\) is \(\sigma\)-relatively-compact
    \item \(F \kmarkwin0 \menGame{X}\)
    \item \(F \markwin \menGame{X}\)
    \item \(F \win \menGame{X}\)
  \end{itemize}
\end{corollary}

\begin{proof}
  Each bullet implies \(X\) is Lindel\"of, so \(X\) may be assumed to be
  regular and second-countable.
\end{proof}

@@

To help describe \(\pl F\kmarkwin2\menGame{X}\) topologically, we introduce a
slight variant of the Menger game and a related covering property.

@@

We may also consider common counterexamples which are finer
than the usual Euclidean line.

@@

\begin{example}
  \(R_{\mathbb Q}\) is non-regular and non-\(\sigma\)-compact, but is
  second-countable and \(\sigma\)-relatively-compact.
\end{example}

\begin{proof}
  Compact sets in \(R_{\mathbb Q}\) cannot contain open intervals,
  and thus are nowhere dense in nonmeager \(\mathbb R\),
  so \(R_{\mathbb Q}\) is not \(\sigma\)-compact. The usual base of
  intervals with rational endpoints (with or without rationals removed)
  witnesses second-countability, and \([-n,n]\) is relatively compact.
\end{proof}

@@

\begin{example}
  \(R_\omega\) is non-regular, non-second-countable, and
  non-\(\sigma\)-relatively-compact, but \(\pl F\win\menGame{R_\omega}\).
\end{example}

\begin{proof}
  If \(S\supseteq\{s_n:n<\omega\}\) for \(s_n\)
  discrete, then \(U_m=R_\omega\setminus\{s_n:m<n<\omega\}\) yields an
  infinite cover \(\{U_m:m<\omega\}\) with no finite subcollection covering \(S\),
  showing that all relatively compact sets are finite, and \(R_\omega\) is not
  \(\sigma\)-relatively-compact.

  Define the winning strategy \(\sigma\) for \(\pl F\) in \(\menGame{R_\omega}\) as
  follows: let \(\sigma(\<\mc{U}_0,\dots,\mc{U}_{2n}\>)=[-n,n]\setminus C_n\)
  for some countable \(C_n=\{c_{n,m}: m<\omega\}\),
  and let \(\sigma(\<\mc{U}_0,\dots,\mc{U}_{2n+1}\>)=\{c_{i,j} : i,j < n\}\).
\end{proof}

We will soon see that, assuming \(\alcompS{\mf c}\), \(\pl F\) has a winning
\(2\)-Mark\"ov strategy for \(\menGame{R_\omega}\) as well.

\begin{proposition}
  If \(X=\bigcup_{i<\omega} X_i\) and \(\pl F \kmarkwin{k} \menGame{X,X_i}\)
  for \(i<\omega\), then \(\pl F \kmarkwin{k} \menGame{X}\)
\end{proposition}

\begin{proof}
  Let \(\sigma_i\) be a \(k\)-Markov strategy
  for \(\pl F\) in \(\menGame{X,X_i}\).

  We define the \(k\)-Markov strategy
  \(\sigma\) for \(\menGame{X}\) as follows:
    \[
      \sigma(\<\mc U_0,\dots,\mc U_n\>)
        =
      \bigcup_{i\leq n}
      \sigma_i(\<\mc U_i,\dots,\mc U_n\>)
    \]

  Let \(\<\mc U_0,\mc U_1,\dots\>\) be a successful counter-attack by \(\pl C\)
  against \(\sigma\). Then there exists \(x\in X_i\) for
  some \(i<\omega\) such that \(x\) is not covered by
  \(\bigcup_{n<\omega}\sigma(\<\mc U_0,\dots,\mc U_n\>)\).
  It follows that \(x\) is not covered by
  \(\bigcup_{n<\omega}\sigma_i(\<\mc U_i,\dots,\mc U_{i+n}\>)\),
  and \(\<\mc U_i,\mc U_{i+1},\dots\>\) is a successful counter-attack
  by \(\pl C\) against \(\sigma_i\).
\end{proof}

\begin{theorem}[\(\alcompS{\mf c}\)]
  \(\pl F\kmarkwin2\menGame{R_\omega}\).
\end{theorem}

\begin{proof}
  It's sufficient to show that \([0,1]\subseteq R_\omega\) is relatively
  robustly Menger. Let \(f_A\) witness \(\alcompS{\mf c}\)
  for \(A\in[[a,b]]^{\leq\omega}\).
  For each open cover \(\mc U\), let \(A_{\mc U}\) be
  such that \([0,1]\setminus A_{\mc U}\) is \(\mc U\)-finite.
  Let \(r_{\mc U}(x)=0\) if \(x\in[0,1]\setminus A_{\mc U}\) and
  \(r_{\mc U}(x)=f_{\mc A_{\mc U}}(x)\) otherwise.
